---
title: "Exponential MLE using Newton-Raphson"
author: "092975 Igobwa Alvin Mugwe, 145054 Kyalo Henry Muoki, 145053 Eunice Mutahi,
106218 Getrude Gichuhi"
date: "3/24/2022"
output: pdf_document
---

```{r}
#tinytex::install_tinytex()
library(tinytex)
```


the exponential pdf is given by $$\begin{equation} f_X(x) = \begin{cases}\lambda \exp(-\lambda x) & x > 0 \\ 0 & \text{otherwise.} \end{cases} \end{equation}$$ 

Finding the likelihood then becomes:$$\begin{equation} \prod_{i=1}^{n} f_X(x;\lambda)=\begin{cases}\prod_{i=1}^{n}\lambda \exp^{(-\lambda x)} & x > 0 \\ 0 & \text{otherwise.}\end{cases}\end{equation}$$

The above equation then opens up to be $$\lambda^{n} \exp(-\lambda*\sum{xi} $$

Introducing logarithms the equation becomes:
$$n\ln{\lambda + -\lambda\sum{xi}} $$

The score function now becomes:

$$\frac{d}{d\lambda} = \frac{n}{\lambda} - \sum{xi}=0 $$
Given that we have now obatined teh score function, we can now use this to create the Newton Raphson but first we need the information function and as we know $$ I(\lambda) = -S'(\lambda)$$


As such $$I(\lambda) = -[-\frac{n}{\lambda^2}] = \frac{n}{\lambda^2}$$


```{r}
exp.mle <- function(x,start,eps=1.e-8,max.iter=100){
  if (missing(start)) start <- median(x)
  lambda <- start
  n <- length(x)
  score <- (n/lambda) - sum(x)
  iter <- 1
  conv <- T
  while (abs(score)>eps && iter<=max.iter){
    info <- n/(lambda**2)
    lambda <- lambda + score/info
    iter <- iter + 1
    score <-( n/lambda) - sum(x)
  }
  if (abs(score)>eps) {
    print("No Convergence")
    conv <- F
  }
  loglik <- n*log(lambda) - lambda*sum(x)
  info <- n/lambda**2
  r <- list(lambda=lambda,loglik=loglik,info=info,convergence=conv)
  r
}


MLE_exp = exp.mle(rexp(100))
MLE_exp
```


Confidence Interval for the lambda obtained
```{r}
upper_bound = MLE_exp$lambda + 1.96*(1/MLE_exp$info)**(0.5)
upper_bound
lower_bound = MLE_exp$lambda - 1.96*(1/MLE_exp$info)**(0.5)
lower_bound
```
$$\text{The confidence bound for the lambda value is found between 1.334121 and 0.8968503 ie} \\
\lambda \approx 1.334121 \pm 0.8968503  $$

