---
title: "Assignment 3"
author: "Getrude Gichuhi"
date: "2022-05 - 14"
output: pdf_document
---

Use the dataset attached to do various extensions of Linear regression. 

Use R Markdown for your submissions. 

Ensure you use sales as the response variable for model fitting. 

Try variety of extensions i.e. Interaction, polynomial, combinations of both interaction and polynomial. 

You can choose any of the covariates. Ensure you interpret your results.


```{r}
#install.packages("readxl")
```
Loading Libraries
```{r}
library(readxl)
library(dplyr)
library(tidyverse)
library(lattice)
```


Load the Dataset 


```{r}
df <- read_excel("Sales_Dataset.xlsx")
print(head(df))
```


```{r cars}
summary(df)
```


```{r}
plot(df)
```
 
 Interaction Models
 
 
```{r}
lm_Sales = lm(Sales~ .,data=df)
summary(lm_Sales)

```
Interaction of TV and Radio 
```{r}
lm_Sales1 = lm(Sales ~ TV* Radio, data=df)
summary(lm_Sales1)
```

interaction of TV and Newspaper
```{r}
lm_Sales2 = lm(Sales ~ TV* Newspaper, data=df)
summary(lm_Sales2)
```

Interaction of TV and Billboard
```{r}
lm_Sales3 = lm(Sales ~ TV* BillBoard, data=df)
summary(lm_Sales3)
```

Interaction of Tv, Billboard, Radio and Newspaper

```{r}
lm_sales4 = lm(Sales ~ TV*BillBoard*Radio*Newspaper, data=df)
summary(lm_sales4)
```

Polynomial
```{r}
scatter.smooth(x=df$TV, y=df$Radio, main="TV & Radio") 
```

```{r}
tv_lm = lm(Sales ~poly(TV,degree=2,raw = TRUE), data=df)

summary(tv_lm)
```
The significance level has a 99.47% chance, which falls close to the empirical rule, but not at the 99.7%.
```{r}
News_lm = lm(Sales ~poly(Newspaper,degree=2,raw = TRUE), data=df)

summary (News_lm)
```
The significance level is at 99.92%. 
```{r}
radio_lm = lm(Sales ~poly(Radio,degree=2,raw = TRUE), data=df)

summary(radio_lm)
```

```{r}
Bill_lm = lm(Sales ~poly(BillBoard,degree=2,raw = TRUE), data=df)

summary(Bill_lm)
```

```{r}
mse_tv<-lm(Sales ~ TV, data=df)

mean(mse_tv$residuals^2) 

#summary(mse_tv)
```

```{r}
mse_radio<-lm(Sales ~ Radio, data=df)

mean(mse_radio$residuals^2) 
```

```{r}
mse_news<-lm(Sales ~ Newspaper, data=df)

mean(mse_news$residuals^2) 
```
```{r}
mse_Bill<-lm(Sales ~ BillBoard, data=df)

mean(mse_Bill$residuals^2) 
```
The best polynomial regression to use is Billboard since it has a small error and the adjusted R-Squared is equal to 1. 




