---
title: "Assignment 1 Linear Models"
author: "Getrude Gichuhi"
date: "28/03/2022"
output: pdf_document
---

Use Salaries as your response variable. Ensure you convert gender and region to factor variables

1) Use R markdown and submit a PDF document

2) Explore the data by doing relevant histograms, boxplots, scatter plots etc and interpret your results

3) Fit a linear regression with salary as response variable and all the rest at exploratory variables

4) Interpret all your output

```{r}
#install.packages("readxl")
```

load the libraries to use

```{r}
library("readxl")
```
loading the dataset

```{r}
df <- read_excel("Data1.xlsx")
print(head(df))
```
```{r}
summary(df)
```


Scatter Plots
```{r}
#plot(df$`Years of Experience`, df$Salaries)

scatter.smooth(x=df$`Years of Experience`, y=df$Salaries, main="Salaries ~ Experience") 
```
Box Plots

```{r}
boxplot(df$Salaries)
```


```{r}
boxplot(df$Salaries,
main = "Salaries and Years of experience",
xlab = "`Years of Experience`",
ylab = "Salaries",
col = "orange",
border = "brown",
horizontal = TRUE,
notch = TRUE
)
```
Histogram

```{r}
Salaries <- df$Salaries
hist(Salaries)
```



```{r}
# histogram with added parameters
hist(Salaries,
main="Maximum Salaries",
xlab="`Years of Experience`",
ylab= "Salaries",
col="darkmagenta",
freq=FALSE
)
```


Salaries increases with the number of years of experience. The more experienced, the more salary is earned. 

```{r}
#plot(df$Age, df$Salaries)

scatter.smooth(x=df$Age, y=df$Salaries, main="Salaries ~ Age")
```

```{r}
boxplot(df$Salaries,
main = "Salaries and Years of experience",
xlab = "Age",
ylab = "Salaries",
col = "black",
border = "blue",
horizontal = TRUE,
notch = TRUE
)
```

```{r}
hist(Salaries,
main="Maximum Salaries",
xlab="Age",
ylab= "Salaries",
col="darkgreen",
freq=FALSE
)
```

From ages 28 to around 33 more people seems to earn more from around 3000 - 3200. The data shows that the more energetic one is and young, the salaries are more, while only a handful of them still earn more as the ages progresses.  
Box Plots


```{r}
library(caret)
library(tidyverse)
library(glmnet)
library(dplyr)
```


```{r}
df <- read_excel("Data1.xlsx")
view(df)
```



```{r}
linearMod <- lm(Salaries ~`Years of Experience` +  Age + Region, data=df)  # build linear regression model on full data
print(linearMod)
```
Based on the outcome it shous that the Intercept of the salaries is = 1711.423 which is Beta zero. while Beta one is 92.792 * Years of Experience, Beta 2 * Age and Beta 3 * Region.





