---
title: "Assignment 2 - Linear Models"
author: "Getrude Gichuhi"
date: '2022-04-08'
output: pdf_document
---
Use the dataset attached to do model selection.

Use R Markdown for your submissions. 

Ensure you change the variable Region to factor variable before model fitting. 

Use the variable name power as your response variable and select the best model using AIC

```{r}
# install.packages("readxl")
```


```{r}
library(readxl)
library(dplyr)
library(tidyverse)
library(lattice)
```

Loading the DataSet
```{r}
df <- read_excel("Dataset2.xlsx")
print(head(df))
```


```{r}
summary(df)
```

Change the Region Variable to the factor variable 

```{r}
df$Region <- as.factor(df$Region)
```

```{r}
summary (df)
```
Build a Model
```{r}
lm_power = lm(Power~ .,data=df)
summary(lm_power)
```

Rates shows no estimates or statistics therefore it's wise to remove it. 

```{r}
df1 = subset(df,select = -c(Rates))

lm_power =  lm(Power~ ., data=df1)
summary(lm_power)
```

Model selection of AIC 
1. Forward Selection 
2. Backward Selection 
3. Mixed Selection 

```{r}
step(lm_power, direction = "forward")
```
The forward stepwise selection shows a model of AIC = -151.08 which includes Power ~ Time + Homes + Region + Sequence. Power as the response Variable.


```{r}
step(lm_power, direction = "backward")
```

The Backward Stepwise selection shows a model of AIC = -152.63 which includes Power ~ Homes + Sequence. 


```{r}
step(lm_power, direction = "both")
```

The mixed selcetion shows an AIC of -152.63. 

The best model to use is the forward selection which has an AIC of -151.08

